---
title: "Climate and Hydrological Scenarios Report for Area of Interest"
output:
  html_document:
    css: style.css
  word_document: default
params:
  area_name: "Default Area"
  plot_data: NULL
  aoi_params: NULL
  aoi_scenarios: NULL
  params_def: NULL
  hydro_plot_data: NULL
  hydro_params: NULL
  hydro_scenarios: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dpi = 300, fig.retina = 2)
library(ggplot2)
library(dplyr)
```
# Report for `r params$area_name`

This report shows the annual time series plots for various climate parameters and scenarios for the selected area of interest.

```{r generate_plots, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=400}
if (is.null(params$aoi_params) || nrow(params$aoi_params) == 0) {
  knitr::asis_output(
    "<p><em>No climate parameters were selected for this report.</em></p>"
  )
} else {
  # Loop through each parameter (e.g., Temperature, Precipitation)
  for (i in seq_len(nrow(params$aoi_params))) {
    param_code <- params$aoi_params$param[i]
    param_name <- params$aoi_params$name[i]

    # --- Data Aggregation for Faceting ---
    all_scenarios_data <- lapply(params$aoi_scenarios, function(scen_code) {
      plot_id <- paste(param_code, scen_code, sep = "_")
      df <- params$plot_data[[plot_id]]
      
      if (is.data.frame(df) && nrow(df) > 0 && !all(is.na(df$value))) {
        df$scenario <- toupper(scen_code)
        return(df)
      }
      return(NULL)
    })
    
    combined_df <- dplyr::bind_rows(all_scenarios_data)

    # --- Header and Plot Generation ---

    # 1. Print the parameter name as a header FIRST.
    # This is the key change: separating the header output from the plot output.
    cat(paste0("<h4>", param_name, "</h4>"))

    # 2. Now, proceed to generate either the plot or the "no data" message.
    if (is.data.frame(combined_df) && nrow(combined_df) > 0) {
      
      param_info <- dplyr::filter(params$params_def, input == param_code)
      y_label <- param_name # Fallback to param_name
      if (nrow(param_info) > 0 && "unit" %in% colnames(param_info)) {
        unit <- param_info$unit
        if (is.character(unit) && !is.na(unit) && nzchar(trimws(unit))) {
          y_label <- unit
        }
      }
      
      if ("date" %in% colnames(combined_df)) {
        combined_df <- combined_df %>%
          mutate(date = as.numeric(format(as.Date(date), "%Y")))
      }

      # title absolute plots
      cat(paste0("<h5>","Absolute values plots", "</h5>"))
      # Create the ggplot object
      line_color <- if (param_code %in% c("pr", "csdi")) "steelblue" else "red"
      line_color_anom <- if (param_code %in% c("pr")) "steelblue" else "red"
      p <- ggplot(combined_df, aes(x = date)) + 
        geom_ribbon(aes(ymin = value10, ymax = value90), fill = "grey70", alpha = 0.5) + 
        geom_line(aes(y = value), color = line_color) + 
        geom_smooth(aes(y = value), method = "lm", se = FALSE, color = "#252525", linewidth = 0.25) + 
        facet_wrap(~scenario, ncol = 2) + 
        scale_x_continuous(breaks = c(1961, 1981,2000, 2014, 2021, 2041,2061, 2081, 2100)) + 
        labs(
          x = "Year",
          y = y_label
        ) + 
        theme_minimal(base_size = 10) + 
        theme(
          strip.text = element_text(size = 10),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 9),
          axis.text.x = element_text(angle = 45, hjust = 1)
        )
      
      # Print the combined plot
      print(p)

      # title anomaly plots
      cat(paste0("<h5>","Anomaly values plots", "</h5>"))
      # --- Anomaly Calculation and Plot ---
      historical_mean <- combined_df %>% 
        filter(date >= 1981 & date <= 2010) %>%
        summarise(mean_value = mean(value, na.rm = TRUE)) %>%
        pull(mean_value)

      if (!is.na(historical_mean)) {
        if (param_code == "pr") {
          if (historical_mean != 0) {
            combined_df <- combined_df %>% 
              mutate(
                anomaly = ((value - historical_mean) / historical_mean) * 100,
                anomaly10 = ((value10 - historical_mean) / historical_mean) * 100,
                anomaly90 = ((value90 - historical_mean) / historical_mean) * 100
              )
            y_label_anomaly <- "Anomaly (%)"
          } else {
            combined_df <- combined_df %>% mutate(anomaly = NA, anomaly10 = NA, anomaly90 = NA)
            y_label_anomaly <- "Anomaly (%)"
          }
        } else {
          combined_df <- combined_df %>% 
            mutate(
              anomaly = value - historical_mean,
              anomaly10 = value10 - historical_mean,
              anomaly90 = value90 - historical_mean
            )
          y_label_anomaly <- paste0("Anomaly (", y_label, ")")
        }

        p_anomaly <- ggplot(combined_df, aes(x = date)) +
          geom_ribbon(aes(ymin = anomaly10, ymax = anomaly90), fill = "grey70", alpha = 0.5) +
          geom_line(aes(y = anomaly), color = line_color_anom) +
          geom_smooth(aes(y = anomaly), method = "lm", se = FALSE, color = "black", linewidth = 0.5) +
          facet_wrap(~scenario, ncol = 2) + 
          scale_x_continuous(breaks = c(1961, 1981,2000, 2014, 2021, 2041,2061, 2081, 2100)) + 
          labs(
            x = "Year",
            y = y_label_anomaly
          ) + 
          theme_minimal(base_size = 10) + 
          theme(
            strip.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 9),
            axis.text.x = element_text(angle = 45, hjust = 1)
          )
        
        print(p_anomaly)
      }

    } else {
      # Fallback message if no data exists for this parameter
      # The header is already printed above, so we only need the message here.
      knitr::asis_output(
        paste0("<p><em>Data not available for this parameter.</em></p>")
      )
    }

    # Add a horizontal rule between parameter sections
    if (i < nrow(params$aoi_params)) {
      knitr::asis_output("\n<hr/>\n\n")
    }
  }
}
```

#### Summarising climate

```{r summary_table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(tidyr)

# Define the periods
hist_period <- 1981:2010
future_periods <- list(
  "2041-2060" = 2041:2060,
  "2061-2080" = 2061:2080,
  "2081-2100" = 2081:2100
)

# Loop through each parameter
for (i in seq_len(nrow(params$aoi_params))) {
  param_code <- params$aoi_params$param[i]
  param_name <- params$aoi_params$name[i]

  # Initialize a list to store summary data for the current parameter
  summary_list <- list()

  # Loop through each scenario
  for (scen_code in params$aoi_scenarios) {
    plot_id <- paste(param_code, scen_code, sep = "_")
    df <- params$plot_data[[plot_id]]
    
    # --- Optional DEBUG (comment out in production) ---
    # knitr::asis_output(paste0("```r\n", paste(capture.output(str(df)), collapse = '\n'), "\n```"))
    # --------------------------------------------------

    if (is.data.frame(df) && nrow(df) > 0) {
      
      # Ensure date is numeric year
      if ("date" %in% colnames(df) && !is.numeric(df$date)) {
        df <- df %>%
          mutate(date = as.numeric(format(as.Date(date), "%Y")))
      }

      # Calculate historical mean
      historical_mean <- df %>%
        filter(date %in% hist_period) %>%
        summarise(mean_value = mean(value, na.rm = TRUE)) %>%
        pull(mean_value)

      if (length(historical_mean) > 0 && !is.na(historical_mean)) {
        
        # Loop through future periods
        for (period_name in names(future_periods)) {
          future_period <- future_periods[[period_name]]
          
          # Calculate future mean
          future_mean <- df %>%
            filter(date %in% future_period) %>%
            summarise(mean_value = mean(value, na.rm = TRUE)) %>%
            pull(mean_value)

          if (length(future_mean) > 0 && !is.na(future_mean)) {
            change <- if (param_code == "pr") {
              if (historical_mean != 0) {
                ((future_mean - historical_mean) / historical_mean) * 100
              } else {
                NA
              }
            } else {
              future_mean - historical_mean
            }
            
            # Add to summary list
            summary_list[[length(summary_list) + 1]] <- data.frame(
              Scenario = toupper(scen_code),
              Period = period_name,
              Historical_Mean = round(historical_mean, 2),
              Future_Mean = round(future_mean, 2),
              Change = round(change, 2)
            )
          }
        }
      }
    }
  }

  # Combine list into a data frame and print table for the current parameter
  if (length(summary_list) > 0) {
    summary_df <- do.call(rbind, summary_list)
    
    # Reshape data to wide format
    summary_wide <- summary_df %>%
      tidyr::pivot_wider(
        names_from = Period,
        values_from = c("Future_Mean", "Change"),
        names_sep = "_"
      )

    # Print title and table
    cat(paste0("<h5>", param_name, "</h5>"))
    print(knitr::kable(
      summary_wide
    ))
  }
}
```

## Hydrological Scenario (RCP)

This section shows annual time series plots for River Discharge under RCP scenarios. The historical baseline period is 1971-2005.

```{r hydro_plots, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=400}
if (is.null(params$hydro_params) || nrow(params$hydro_params) == 0) {
  knitr::asis_output(
    "<p><em>No hydrological parameters were selected for this report.</em></p>"
  )
} else {
  # Loop through each hydro parameter
  for (i in seq_len(nrow(params$hydro_params))) {
    param_code <- params$hydro_params$param[i]
    param_name <- params$hydro_params$name[i]

    # Data Aggregation for Faceting
    all_hydro_data <- lapply(params$hydro_scenarios, function(scen_code) {
      plot_id <- paste(param_code, scen_code, sep = "_")
      df <- params$hydro_plot_data[[plot_id]]
      
      if (is.data.frame(df) && nrow(df) > 0 && !all(is.na(df$value))) {
        # Format RCP label nicely
        scen_label <- gsub("_", " ", toupper(scen_code))
        df$scenario <- scen_label
        return(df)
      }
      return(NULL)
    })
    
    combined_hydro_df <- dplyr::bind_rows(all_hydro_data)

    # Header and Plot Generation
    cat(paste0("<h4>", param_name, "</h4>"))

    if (is.data.frame(combined_hydro_df) && nrow(combined_hydro_df) > 0) {
      
      y_label <- "m³/s"
      
      if ("date" %in% colnames(combined_hydro_df)) {
        combined_hydro_df <- combined_hydro_df %>%
          mutate(date = as.numeric(format(as.Date(date), "%Y")))
      }

      # Absolute values plots
      cat(paste0("<h5>", "Absolute values plots", "</h5>"))
      line_color <- "steelblue"
      p <- ggplot(combined_hydro_df, aes(x = date)) + 
        geom_ribbon(aes(ymin = value10, ymax = value90), fill = "grey70", alpha = 0.5) + 
        geom_line(aes(y = value), color = line_color) + 
        geom_smooth(aes(y = value), method = "lm", se = FALSE, color = "#252525", linewidth = 0.25) + 
        facet_wrap(~scenario, ncol = 3) + 
        scale_x_continuous(breaks = c(1971, 1990, 2010, 2030, 2050, 2070, 2099)) + 
        labs(
          x = "Year",
          y = y_label
        ) + 
        theme_minimal(base_size = 10) + 
        theme(
          strip.text = element_text(size = 10),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 9),
          axis.text.x = element_text(angle = 45, hjust = 1)
        )
      
      print(p)

      # Anomaly plots (using 1971-2005 baseline)
      cat(paste0("<h5>", "Anomaly values plots (% change vs 1971-2005)", "</h5>"))
      historical_mean <- combined_hydro_df %>% 
        filter(date >= 1971 & date <= 2005) %>%
        summarise(mean_value = mean(value, na.rm = TRUE)) %>%
        pull(mean_value)

      if (!is.na(historical_mean) && historical_mean != 0) {
        combined_hydro_df <- combined_hydro_df %>% 
          mutate(
            anomaly = ((value - historical_mean) / historical_mean) * 100,
            anomaly10 = ((value10 - historical_mean) / historical_mean) * 100,
            anomaly90 = ((value90 - historical_mean) / historical_mean) * 100
          )
        y_label_anomaly <- "Anomaly (%)"

        p_anomaly <- ggplot(combined_hydro_df, aes(x = date)) +
          geom_ribbon(aes(ymin = anomaly10, ymax = anomaly90), fill = "grey70", alpha = 0.5) +
          geom_line(aes(y = anomaly), color = "steelblue") +
          geom_smooth(aes(y = anomaly), method = "lm", se = FALSE, color = "black", linewidth = 0.5) +
          facet_wrap(~scenario, ncol = 3) + 
          scale_x_continuous(breaks = c(1971, 1990, 2010, 2030, 2050, 2070, 2099)) + 
          labs(
            x = "Year",
            y = y_label_anomaly
          ) + 
          theme_minimal(base_size = 10) + 
          theme(
            strip.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 9),
            axis.text.x = element_text(angle = 45, hjust = 1)
          )
        
        print(p_anomaly)
      }

    } else {
      knitr::asis_output(
        paste0("<p><em>Data not available for this parameter.</em></p>")
      )
    }

    if (i < nrow(params$hydro_params)) {
      knitr::asis_output("\n<hr/>\n\n")
    }
  }
}
```

#### Summarising hydrological scenario

```{r hydro_summary_table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Define the periods for hydro (historical: 1971-2005)
hydro_hist_period <- 1971:2005
hydro_future_periods <- list(
  "2041-2060" = 2041:2060,
  "2061-2080" = 2061:2080,
  "2081-2099" = 2081:2099
)

if (!is.null(params$hydro_params) && nrow(params$hydro_params) > 0) {
  for (i in seq_len(nrow(params$hydro_params))) {
    param_code <- params$hydro_params$param[i]
    param_name <- params$hydro_params$name[i]

    summary_list <- list()

    for (scen_code in params$hydro_scenarios) {
      plot_id <- paste(param_code, scen_code, sep = "_")
      df <- params$hydro_plot_data[[plot_id]]

      if (is.data.frame(df) && nrow(df) > 0) {
        if ("date" %in% colnames(df) && !is.numeric(df$date)) {
          df <- df %>%
            mutate(date = as.numeric(format(as.Date(date), "%Y")))
        }

        historical_mean <- df %>%
          filter(date %in% hydro_hist_period) %>%
          summarise(mean_value = mean(value, na.rm = TRUE)) %>%
          pull(mean_value)

        if (length(historical_mean) > 0 && !is.na(historical_mean) && historical_mean != 0) {
          for (period_name in names(hydro_future_periods)) {
            future_period <- hydro_future_periods[[period_name]]
            
            future_mean <- df %>%
              filter(date %in% future_period) %>%
              summarise(mean_value = mean(value, na.rm = TRUE)) %>%
              pull(mean_value)

            if (length(future_mean) > 0 && !is.na(future_mean)) {
              # Percentage change for discharge
              change <- ((future_mean - historical_mean) / historical_mean) * 100
              scen_label <- gsub("_", " ", toupper(scen_code))
              
              summary_list[[length(summary_list) + 1]] <- data.frame(
                Scenario = scen_label,
                Period = period_name,
                Historical_Mean = round(historical_mean, 1),
                Future_Mean = round(future_mean, 1),
                Change_Pct = round(change, 1)
              )
            }
          }
        }
      }
    }

    if (length(summary_list) > 0) {
      summary_df <- do.call(rbind, summary_list)
      
      summary_wide <- summary_df %>%
        tidyr::pivot_wider(
          names_from = Period,
          values_from = c("Future_Mean", "Change_Pct"),
          names_sep = "_"
        )

      cat(paste0("<h5>", param_name, "</h5>"))
      print(knitr::kable(
        summary_wide
      ))
    }
  }
}
```
